Commutative Character Sheaves

### 2015-09-11 followup.tex
Friday catchup
hey!

\title[Commutative Character Sheaves]{Commutative Character Sheaves, with an application to generic types for supercuspidal representations}
\author{Clifton Cunningham}
\address{Department of Mathematics and Statistics, University of Calgary, 2500 University Drive Northwest, Calgary, AB, Canada, {T2N~1N4}.}
\email{cunning@math.ucalgary.ca}
\author{David Roe}
\address{Department of Mathematics, University of Pittsburgh, 301 Thackeray Hall, Pittsburgh, PA , United States, 15260.}
\email{roed.math@gmail.com}

@@

\section{Application to certain supercuspidal representations of $p$-adic groups}

[A basic question which I think we must answer but which we can maybe skate around for now:
When is the Suzuki transform of a K-variety representable as a k-scheme?

@@

@@

\section{Descent along smooth fibrations}

Let $G$ be a group scheme acting on $X$ by $m : G \times X \to X$.

A principal $G$-bundle $f : X\to Y$ is a coequalizer of $m$ and projection $G\times X \to X$  
which is locally trivial in the following sense.
First, there is an open covering $\{ V_n\}$ of $Y$ such that, for each $V$ in this covering, the lower coequalizer diagram is a base change of the one above
\[
\begin{tikzcd}
G\times X
 \arrow{rr}[shift left]{m}
  \arrow{rr}[shift right,swap]{\text{proj'n}}
&& X \arrow{r}{f} &  Y\\
G\times  (G\times V) \arrow{u}
 \arrow{rr}[shift left]{\text{mult'n in $G$}}
  \arrow{rr}[shift right,swap]{\text{proj'n}}
&& G\times V \arrow{u}  \arrow{r} &  \arrow{u} V;
\end{tikzcd}
\]

@@

\begin{proposition}
Let $G$, $X$ and $Y$ be smooth.
Let $f : X \to Y$ be a $G$-torsor.
Then the comparison functor, pictured below,
\[
\begin{tikzcd}
%\Loc^1(G \times X)
%& \arrow{l}[shift left,swap]{m^*}
%  \arrow{l}[shift right]{p^*}
\Loc^1(X) 
&&\arrow{ll}[swap]{f^*} 
\Loc^1(Y) \arrow{dl}{\text{comparison}}\\
& \arrow{ul}{\text{forget}} \Loc^1_G(X) & 
\end{tikzcd}
\]
is an equivalence.
\end{proposition}

@@

It remains to be shown that $\mathcal{G}$ is a perverse sheaf. Again, we work locally. For each such $n$,
\begin{eqnarray*}
{f}_n^*[\dim{P}] \mathcal{F}'_n 
	&=& {f}_n^*[\dim{P}] {e}_n^* \mathcal{F}_n [-\dim{P}]\\
	&\iso& ({e}_n\circ {f}_n)^* \mathcal{F}_n.
\end{eqnarray*}
Let ${u}_n : {X}_n \to {P}\times{X}_n$ be the section of ${m}_n : {P}\times{X}_n \to {X}_n$ corresponding to $1$ (so ${u}_n$ is the unique morphism of varieties such that $(\id\times{t}_n)\circ {u}_n \circ {t}_n^{-1} (h,y) = (h,1,y)$). The domain of ${u}_n^*\mu_{\mathcal{F}_n}$ is ${u}_n^*{m}_n^* \mathcal{F}_n \iso ({m}_n \circ {u}_n)^* \mathcal{F}_n = \mathcal{F}_n$, since ${u}_n$ is a section of ${m}_n$; the codomain of ${u}_n^* \mu_{\mathcal{F}_n}$ is ${u}_n^* p^* \mathcal{F}_n \iso (p\circ {u}_n)^*\mathcal{F}_n$. Since $p\circ {u}_n = {e}_n \circ {f}_n$, it follows that 
\[
{u}_n^* \mu_{\mathcal{F}_n} : \mathcal{F}_n \to {f}_n^*[\dim{P}] \mathcal{F}'_n
\]
is an isomorphism in $D^b_c({X}_n, \EE)$. By \cite[Prop~4.2.5]{BBD} and the fact that ${\mathcal M}{X}_n$ is stable in $D^b_c({X}_n, \EE)$ under isomorphisms, it follows that $\mathcal{F}'_n \in \obj {\mathcal M}{Y}_n$. By standard glueing arguments, the collection of isomorphisms ${u}_n^* \mu_{\mathcal{F}_n} \in \mor{\mathcal M}{X}_n$, as ${X}_n$ ranges over the open cover of ${X}$ fixed above, uniquely determines an isomorphism in \[
\Hom_{D^b_c({X}, \EE)}(\mathcal{F},{f}^*[\dim{P}] \mathcal{G}).
\]  
From \cite[Prop~4.2.5]{BBD}  it follows that $\mathcal{G}\in \obj{\mathcal M}{Y}$. This completes the proof of Proposition~\ref{proposition: torsor}.
\end{proof}

>>>>

\title[Commutative Character Sheaves]{Commutative Character Sheaves}
%with an application to geometric types for supercuspidal representations
\author{Clifton Cunningham}
\address{Department of Mathematics and Statistics, University of Calgary, 2500 University Drive Northwest, Calgary, AB, Canada, {T2N~1N4}.}
\email{cunning@math.ucalgary.ca}
\author{David Roe}
\address{Department of Mathematics, University of Pittsburgh, 301 Thackeray Hall, Pittsburgh, PA , United States, 15260.}
\email{roed.math@gmail.com}

@@

\section{Application: geometric types for supercuspidal representations}

@@

{Application to certain supercuspidal representations of $p$-adic groups}

@@

\section{Descent along torsors}

Let $G$ be a group scheme acting on $X$ by $m : G \times X \to X$; let $p : G\times X \to X$ be projection.

Recall that a morphism $f : X\to S$ is a regular epimorphism if . . . . coequalizer for the pair of morphisms $(m,p)$ if ....

A morphism $f : X\to S$ is a $G$-torsor if it is equivariant for the trivial action of $G$ on $S$ and if there is a covering $\{ S_i \tq i \in I\}$ of $S$ such that, for each $i\in I$, there is an isomorphism $X\times_S S_i \to G\times S_i$ such that composition with projection $G\times S_i \to S_i$ is $X\times_S S_i \to S_i$.

We recall the definition of the category $\Loc_G(X)$ of $G$-equivariant local systems on $X$.
Objects in $\Loc_G(X)$ are pairs $(\cs{L}, \nu)$ where $\cs{L} \in \Loc(X)$ and $\nu : m^*\cs{L} \to p^*\cs{L}$ is an isomorphism in $\Loc(G\times X)$
such that the following diagram commutes, where $m_0(a,b,x) = (a,m(b,x))$ and $m_1(a,b,x) = (m(ab),x)$ and $p_2(a,b,x) = (b,x)$.
\[
\begin{tikzcd}
\ &  m_0^*  m^*\, \mathcal{L} \arrow{r}{m_0^*\nu} & m_0^* p^*\, \mathcal{L} & \\
m_1^* m^*\, \mathcal{L} \arrow{dr}[swap]{m_1^*\nu} \arrow[equal]{ur} &&& \arrow[equal]{ul}  \arrow{dl}{p_2^*\nu} p_2^* m^*\, \mathcal{L} \\
& m_1^* p^*\, \mathcal{L} \arrow[equal]{r} & p_2^* p^*\, \mathcal{L} & 
\end{tikzcd}
\]

A morphism $(\cs{L},\nu) \to (\cs{L}',\nu')$ in $\Loc_G(X)$ is a morphism $\alpha: \cs{L} \to \cs{L}'$ in $\Loc(X)$ such that . . . .

Let $f : X\to S$ be a regular epimorphism with kernel pair $(m,p)$. 
If $\cs{S} \in \Loc(S)$ then $f^*\cs{S} \in \Loc(X)$ comes equipped with a canonical isomorphism $m^* (f^* \cs{S}) \to p^* (f^*\cs{S})$, since $f\circ m = f\circ p$.
This defines $f_G^* : \Loc(S) \to \Loc_G(X)$ such that the functor $f^* : \Loc(S) \to \Loc(X)$ factors as below.
\[
\begin{tikzcd}
\Loc(X) 
&&\arrow{ll}[swap]{f^*} 
\Loc(S) \arrow{dl}{f_G^*}\\
& \arrow{ul}{\text{forget}} \Loc_G(X) & 
\end{tikzcd}
\]

\begin{lemma}\label{lem:descent}
If $f: X\to S$ is a $G$-torsor and a regular epimorphism for $(m,p)$, then $f_G^* : \Loc(S) \to \Loc_G(X)$ admits a left-adjoint.
% $f^G_! : \Loc_G(X) \to \Loc(S)$
\end{lemma}

\begin{remark}
Comparison with the comparison functor
\end{remark}

\begin{lemma}
In this case, the functor $f^* : \Loc(S)\to \Loc(X)$ admits a left-adjoint 
\end{lemma}
\begin{proof}
The left-adjoint $f_! : \Loc(X) \to \Loc(S)$ is defined as follows.
Now, suppose $\cs{L}\in \Loc(X)$.
For each $i\in I$, define $\cs{S}_i \in \Loc(S_i)$ by $\cs{S}_i\ceq t_i^* \cs{L}$.
The sheaves $\cs{S}_i$ glue to define a sheaf $\cs{S}$ on $S$ by definition above.
Morphisms of these sheaves are likewise defined locally . . .

\end{proof}

is a coequalizer of $m$ and projection $G\times X \to X$  
which is locally trivial in the following sense.
First, there is an open covering $\{ V_n\}$ of $Y$ such that, for each $V$ in this covering, the lower coequalizer diagram is a base change of the one above

@@

@@

It remains to be shown that $\mathcal{G}$ is a perverse sheaf. Again, we work locally. For each such $n$,
\begin{eqnarray*}
{f}_n^*[\dim{P}] \mathcal{F}'_n 
	&=& {f}_n^*[\dim{P}] {e}_n^* \mathcal{F}_n [-\dim{P}]\\
	&\iso& ({e}_n\circ {f}_n)^* \mathcal{F}_n.
\end{eqnarray*}
Let ${u}_n : {X}_n \to {P}\times{X}_n$ be the section of ${m}_n : {P}\times{X}_n \to {X}_n$ corresponding to $1$ (so ${u}_n$ is the unique morphism of varieties such that $(\id\times{t}_n)\circ {u}_n \circ {t}_n^{-1} (h,y) = (h,1,y)$). The domain of ${u}_n^*\mu_{\mathcal{F}_n}$ is ${u}_n^*{m}_n^* \mathcal{F}_n \iso ({m}_n \circ {u}_n)^* \mathcal{F}_n = \mathcal{F}_n$, since ${u}_n$ is a section of ${m}_n$; the codomain of ${u}_n^* \mu_{\mathcal{F}_n}$ is ${u}_n^* p^* \mathcal{F}_n \iso (p\circ {u}_n)^*\mathcal{F}_n$. Since $p\circ {u}_n = {e}_n \circ {f}_n$, it follows that 
\[
{u}_n^* \mu_{\mathcal{F}_n} : \mathcal{F}_n \to {f}_n^*[\dim{P}] \mathcal{F}'_n
\]
is an isomorphism in $D^b_c({X}_n, \EE)$. By \cite[Prop~4.2.5]{BBD} and the fact that ${\mathcal M}{X}_n$ is stable in $D^b_c({X}_n, \EE)$ under isomorphisms, it follows that $\mathcal{F}'_n \in \obj {\mathcal M}{Y}_n$. By standard glueing arguments, the collection of isomorphisms ${u}_n^* \mu_{\mathcal{F}_n} \in \mor{\mathcal M}{X}_n$, as ${X}_n$ ranges over the open cover of ${X}$ fixed above, uniquely determines an isomorphism in \[
\Hom_{D^b_c({X}, \EE)}(\mathcal{F},{f}^*[\dim{P}] \mathcal{G}).
\]  
From \cite[Prop~4.2.5]{BBD}  it follows that $\mathcal{G}\in \obj{\mathcal M}{Y}$. This completes the proof of Proposition~\ref{lem:descent}.
\end{proof}

\begin{proposition}
Let $f : X \to S$ be a morphism of smooth group schemes.
Suppose $G$ is a smooth group scheme acting on $X$.
Suppose $f$ is a $G$-torsor.
Then the comparison functor, pictured below,
\[
\begin{tikzcd}
\CCS(X) 
&&\arrow{ll}[swap]{f^*} 
\CCS(S) \arrow{dl}{\text{comparison}}\\
& \arrow{ul}{\text{forget}} \CCS_G(X) & 
\end{tikzcd}
\]
is an equivalence.
\end{proposition}
